---
title: "Stops"
author: "Faisal Jan"
format: html
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(gt)
library(marginaleffects)
```

```{r}
#| cache: true
x <- stops |>
  filter(race %in% c("black", "white")) |>
  mutate(race = str_to_title(race), sex = str_to_title(sex))

fit_stops <- linear_reg() |>
  set_engine("lm") |>
  fit(arrested ~ sex + race * zone, data = x)
```

```{r}
# Plot
plot_predictions(fit_stops, condition = c("sex", "race"), draw = FALSE) |>
  mutate(
    group = paste(sex, race),
    group = factor(
      group,
      levels = c("Female Black", "Female White", "Male Black", "Male White")
    )
  ) |>
  ggplot(aes(x = group, y = estimate, fill = race)) +
  geom_col(width = 0.6, color = "black") +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  labs(
    title = "Predicted Probability of Arrest by Race and Sex",
    subtitle = "Black drivers—especially men—face higher predicted arrest rates during traffic stops",
    caption = "Source: Stanford Open Policing Project (New Orleans Data)",
    x = "Driver Group",
    y = "Predicted Probability of Arrest"
  ) +
  scale_fill_manual(values = c("Black" = "#1f77b4", "White" = "#ff7f0e")) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12, margin = margin(b = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10)),
    legend.position = "none"
  )
```

Fair and unbiased decision-making in law enforcement is crucial, especially when outcomes like arrests may be influenced by personal characteristics such as race. Using data from about 400,000 traffic stops in New Orleans and compiled by the Stanford Open Policing Project, we seek to examine the difference in arrest rates between Black and White drivers vary by location or time of day. This sample of entries from the real data, which is nearly 3.1 million records, may not fully represent the population. In Zone W, the predicted probability of arrest for Black male drivers is approximately 35.3% (95% CI: 33.5% to 37.1%), compared to just 28.1% (95% CI: 26.3% to 29.9%) for White male drivers, making it the zone with the highest overall arrest likelihood and one of the largest racial disparities.

$$
P(Y = 1) = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n)}}
$$

with $Y \sim \text{Bernoulli}(\rho)$ where $\rho = \frac{1}{1 + e^{-(\beta_0 + \beta_1 X_1 + \cdots + \beta_n X_n)}}$

<!-- Fitted Model -->
\begin{align*}
\hat{\text{arrested}} =\ & 0.177 
+ 0.0614 \cdot \text{sexMale}
- 0.0445 \cdot \text{raceWhite} \\
&+ 0.0146 \cdot \text{zoneB}
+ 0.00610 \cdot \text{zoneC}
+ 0.0781 \cdot \text{zoneD} \\
&+ 0.00190 \cdot \text{zoneE}
- 0.00271 \cdot \text{zoneF}
+ 0.0309 \cdot \text{zoneG}
+ 0.0757 \cdot \text{zoneH}
\end{align*}

```{r}
model_tidy <- fit_stops |> tidy(conf.int = TRUE)

# Generate the LaTeX table
model_tidy |>
  select(term, estimate, conf.low, conf.high) |>
  mutate(across(where(is.numeric), ~ round(.x, 4))) |>
  gt() |>
  cols_label(
    term = "Variable",
    estimate = "Estimate",
    conf.low = "95% CI Lower",
    conf.high = "95% CI Upper"
  ) |>
  tab_header(
    title = "Regression Estimates for Arrest Model"
  ) |>
  fmt_number(
    columns = c(estimate, conf.low, conf.high),
    decimals = 4
  ) |>
  opt_table_font(
    font = list(gt::google_font("Lato"), "sans-serif")
  )
```
